A forest is more than its trees: haplotypes and inferred ARGs

### 2024-10-22 paper/writeup.tex
more writeup, sample figure with "correct dpi"
It doesn't work

\begin{figure}[!ht]
\begin{center}
	\includegraphics[width=5in]{edge_extend_method.pdf}
\end{center}
\caption{A visualization of the \textit{extend haplotypes} method. Given two neighboring trees $T_1$ and $T_2$, there exists paths from root node $6$ to sample node $0$. $T_1$ contains the path $6\to 4\to 0$ while $T_2$ has path $6\to 5\to 0$. The intermediate nodes $4$ and $5$ do not appear in $T_2$ respectively $T_1$, and so we consider these two paths \textit{mergeable}. The extend haplotypes method joins these two paths, constructing the merged path $6\to 5\to 4\to 0$ for both $T_1$ and $T_2$.
    \label{fig:extending_diagram}
}
\end{figure}

@@

A relatively efficient algorithm to do this is described in \algorithmref{alg:extend}, 
and implemented in \tskit{} as \texttt{extend\_haplotypes}.
The algorithm considers each tree transition from $T_k$ to $T_{k+1}$ in turn, updating its internal state
(which includes possibly modifying $T_k$) as it goes.
Suppose we are at the transition from tree $T_k$ to tree $T_{k+1}$,
which is done by first removing a set of edges $O$
and then adding another set of edges $I$.
$O$ defines a sub-forest $F_O$ of $T_k$,
and $I$ defines a sub-forest $F_I$ of $T_{k+1}$.
The key step in the algorithm is to determine whether the pair of paths
that terminate in a given node in two adjacent trees are mergeable,
\algorithmref{alg:mergeable}, which works as follows.
If a pair of paths is mergeable,
then the edges of the two paths must lie in $O$ and $I$, respectively.
Suppose an edge in $O$ has child $c$.
To see if $c$ is the base of a pair of mergeable paths,
the algorithm traverses up from $c$ in both $F_O$ and $F_I$;
terminating if a node in the other tree is found
(i.e., if the node traversed in $F_O$ is in $T_{k+1}$ or if the node traversed in $F_I$ is in $T_k$)
or if a pair of traversed nodes have the same time.
If these two traversals end in the same node $p$, the paths are mergeable.
Iterating over all edges in $O$ will thus find all mergeable pairs of paths.
There is often more than one pair of mergeable paths in a tree transition;
so, the algorithm merges pairs of mergeable paths,
starting with pairs that add the smallest number of new edges,
until no more are found.
\comment{Maybe we need to highlight in a figure how new edges can be added.}

\algorithmref{alg:extend} simplifies the algorithm in several ways --
for instance, the bookeeping required to keep track of $T_k$ and $T_{k+1}$ is omitted.
Furthermore, as described the algorithm does one left-to-right pass over the tree sequence;
in practice we do repeated passes in both directions until no changes can be made.
The main step that is omitted is a description of the \texttt{merge} operation,
which performs the actual extending of haplotypes.
This algorithm is essentially the same as \texttt{mergeable} in \algorithmref{alg:mergeable},
except with additional bookkeeping.
Roughly speaking, the algorithm
traverses up from the shared base node $c$,
doing the appropriate operations to insert the nodes along the path found in $T_k$
into the path in $T_{k+1}$.
To do this, some edges that end at $a_k$ will be extended to end at $a_{k+1}$;
some edges that begin at $a_k$ will be postponed to begin at $a_{k+1}$,
and some entirely new edges may be added.
Furthermore, the trees $T_k$ and $T_{k+1}$ (and corresponding forests $F_O$ and $F_I$)
need to be updated.

@@

@@

 \begin{algorithm2e}[!ht]
 \SetStartEndCondition{ }{}{}%
 \SetKwProg{Fn}{def}{\string:}{}
 \SetKwFunction{Range}{range}%%
 \SetKw{KwTo}{in}
 \SetKwFor{For}{for}{\string:}{}%
 \SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
 \SetKwFor{While}{while}{:}{fintq}%
 \SetKw{Break}{break}
 \newcommand\forcond{$i$ \KwTo\Range{$n$}}
 \SetKwComment{Comment}{}
 \AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
 \DontPrintSemicolon\LinesNotNumbered

 \caption{Discrepancy Function}\label{alg:disc}
 \input{$(T_1,T_2)$}
 \output{tree discrepancy, root-mean-square error}

 \Fn{Shared Spans($T_1,T_2$)}{
 \For{pairs $(i,j)\in N_{T_1}\times N_{T_2}$}{
 	$i = (t_i, L_i)$ and $j = (t_j, L_j)$ where $L_i, L_j\subset \left[0,L\right).\quad$ \Comment{Each node is a tuple (time, genome location).}
 	\If{$(i,j)$ has equal sample sets, $s(i)=s(j)$}{
 		Define $a_{ij}=|L_i\cap L_j|$.
 	}
 	\Else{Define $a_{ij}=0$.}
 Define matrix $A=\left[a_{ij}\right]\in M_{N_{T_1}\times N_{T_2}}(\R)$.\;
 }
 \KwRet{$A$}
 }
\BlankLine
\Fn{Discrepancy ($T_1,T_2$)}{
\For{each node $i\in N_{T_1}$}{
	Find nodes in $N_{T_2}$ which have the highest matching span.\;
	$\operatorname{matches}_i = \argmax_{k\in N_{T_2}} a_{ik}.$\;
	Let $j = \argmax_{k\in \operatorname{matches}_i}\frac{1}{1+|t_i-t_k|}$.\;
	Define $\alpha_i = a_{ij}$.\;
	\BlankLine
}
$$tree\ discrepancy = 1 - \frac{\sum_{i\in N_{T_1}} \alpha_i}{\operatorname{tr}\left(Shared Spans(T_1, T_1)\right)}$$\;
\BlankLine\BlankLine
$$rmse = \sqrt{\frac{\sum_{i\in N_{T_1}}|t_i - t_j|\alpha_i}{\operatorname{tr}\left(Shared Spans(T_1,T_1)\right)}}$$\;
\BlankLine
$$true\ proportion = \frac{\sum_{i\in N_{T_1}} \alpha_i}{\operatorname(tr)\left( Shared Spans(T_2, T_2)\right)}$$\;
\BlankLine\BlankLine
\KwRet{tree discrepancy, rmse, true proportion}
}
\end{algorithm2e}

@@

\begin{figure}[bht]
	\includegraphics[width=0.9\linewidth]{newplots_wo_ee/node_spans_wo_trapped_unary_regions.pdf}
	\caption{
        Node spans of simulated tree sequence (blue), simplified tree sequence (green), and extended tree sequence (red).
        \comment{TODO: nodes should be in the same order in all three lines,
        using the ordering from the true tree sequence.}
        \comment{TODO: add plot with percent wrongly added by node?}
    }
	\label{fig:node-spans}
\end{figure}

@@

@@

\begin{algorithm2e}[!ht]
	\SetStartEndCondition{ }{}{}%
	\SetKwProg{Fn}{def}{\string:}{}
	\SetKwFunction{Range}{range}%%
	\SetKw{KwTo}{in}
	\SetKwFor{For}{for}{\string:}{}%
	\SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
	\SetKwFor{While}{while}{:}{fintq}%
	\SetKw{Break}{break}
	\newcommand\forcond{$i$ \KwTo\Range{$n$}}
	\AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%

	\caption{Extends haplotypes across a tree sequence $TS$.
    \comment{I think we can remove this now?}}
	\label{alg:hap}

	\SetKwFunction{ex}{Extend Paths}
	\SetKwFunction{fp}{Find Paths}
	\DontPrintSemicolon
	\LinesNotNumbered
	%\SetAlgoNoEnd
	\input{A tree sequence $\T$}
	\output{New tree sequence,$\T'$}
	\BlankLine
	\Fn{\fp (tree sequence $\T$, forwards = \tn{True})}{
	{
		\eIf{forwards=\tn{True}}{$\T = \left(T_1,...,T_{|\T|}\right)$\BlankLine}
		{\BlankLine
			$\T = \left(T_{|\T|},..., T_1\right)$\;
		}
	}
		\For{pairs $(T_k, T_{k+1})\in \T\times\T$}{
		Compute subforests $F_O\subseteq T_k$ and $F_I\subseteq\in T_{k+1}$.\;
		\For{edges $e$ in $F_O$}{
			Let $c$ be child node.\;
			$p_c^{F_O} = [c]$\;
			$p_c^{F_I} = [c]$\;
			$\bar{p}_c = []$ Path to be inserted\;
			\While{$\exists f\in F_O$ such that $\tn{child}(f)=c$ and $\tn{deg}(\tn{parent}(f))=0$}{
				$p = \tn{parent}(f)$\;
				Append $p$ to $p_c^{F_O}$\;
				$c = p$\;
			}
			\While{$\exists f\in F_I$ such that $\tn{child}(f)=c$ and $\tn{deg}(\tn{parent}(f))<3$}{
				$p =\tn{parent}(f)$\;
				Append $p$ to $p_c^{F_I}$\;
				$c = p$\;
			}
			\If{($p_c^{F_O}\cap p_c^{F_I})\backslash\{c\}\neq\emptyset$}{
				Merge paths up to intersection point\;
				$\bar{p}_c = \hat{p}_c^{F_I} \cup \hat{p}^{F_I}_c$\;
				and sort by node time.\;
			}
			Insert $\bar{p}_c$ into $T_{k+1}$\;		
		}
	}\;
	Update $\T$.\;
	\KwRet{$\T$}\;
}
\BlankLine						
\Fn{\ex ($\T$)}{
	$\T' \gets $ \fp($\T$, forwards =\tn{True})\;
	$\T' \gets $ \fp($\T$, forwards = \tn{False})\;
	\While{Number of edges in $\T$ $\neq$ Number of edges in $\T'$}{
		$\T' \gets $ \fp($\T$, fowards=\tn{True})\;
		$\T' \gets $ \fp($\T$, fowards=\tn{False})\;

	}
	\KwRet{$\T'$}\;
}
\end{algorithm2e}

% Can most likely delete the extend-edges algorithm %%
\begin{algorithm2e}[!ht]

\SetStartEndCondition{ }{}{}%
\SetKwProg{Fn}{def}{\string:}{}
\SetKwFunction{Range}{range}%%
\SetKw{KwTo}{in}
\SetKwFor{For}{for}{\string:}{}%
\SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
\SetKwFor{While}{while}{:}{fintq}%
\SetKw{Break}{break}
\newcommand\forcond{$i$ \KwTo\Range{$n$}}
\AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%

\caption{Extends edges across a tree sequence $TS$. 
\comment{Can be removed, probably -- unless this seems a lot simpler?}}
\label{alg:edge}

\SetKwFunction{ex}{Extend Edges}
\SetKwFunction{fwd}{Extend}
\DontPrintSemicolon
\LinesNotNumbered
%\SetAlgoNoEnd
\input{A tree sequence $\T$}
\output{New tree sequence,$\T'$}

\BlankLine

\Fn{\fwd (tree sequence $\T$, forwards = \tn{True})}{
{	
\eIf{forwards=\tn{True}}{$\T = \left(T_1,...,T_{|\T|}\right)$\BlankLine}
{\BlankLine
	$\T = \left(T_{|\T|},..., T_1\right)$\;}
}
	\For{$T_k\in \T$}{
	Find all edge pairs $(e_1,e_2)$ such that $e_1 = a\to b$ and $e_2 = b\to c$ for some nodes $a,b,c$.\;
	\For{edges $f$ in $T_{k+1}$}{
		\If{$f = a \to c$ and $b$ is not in $T_{k+1}$}{
		Remove $f$ from $T_{k+1}$.\;
		Add $e_1$ and $e_2$ to $T_{k+1}$.\;
		Update $\T$.\;
	}
	}
}\;						
\KwRet{$\T$}\;
}
\BlankLine						
\Fn{\ex ($\T$, iter=100)}{
	\For{$*$ \KwTo \Range(iter)}{
		$\T' \gets \T$.\fwd(fowards=\tn{True})\;
		$\T' \gets \T$.\fwd(fowards=\tn{False})\;
		\If{Number of edges in $\T$ = Number of edges in $\T'$}{
			\Break\BlankLine}
		}
}
\KwRet{$\T'$}\;
\end{algorithm2e}

>>>>

\begin{figure}[!ht]
\begin{center}
	\includegraphics[width=5in]{edge_extend_method.pdf}
\end{center}
\caption{A visualization of the \textit{extend haplotypes} method. Given two neighboring trees $T_1$ and $T_2$, there exists paths from root node $6$ to sample node $0$. $T_1$ contains the path $6\to 4\to 0$ while $T_2$ has path $6\to 5\to 0$. The intermediate nodes $4$ and $5$ do not appear in $T_2$ respectively $T_1$, and so we consider these two paths \textit{mergeable}. The extend haplotypes method joins these two paths, constructing the merged path $6\to 5\to 4\to 0$ for both $T_1$ and $T_2$. This new path extends the edge $4\to 0$ to $T_2$ as well as constructs a new edge $5\to 4$ in the tree sequence.
    \label{fig:extending_diagram}
}
\end{figure}

@@

A relatively efficient algorithm to do this is described in \algorithmref{alg:extend}, 
and implemented in \tskit{} as \texttt{extend\_haplotypes}.
The algorithm considers each tree transition from $T_k$ to $T_{k+1}$ in turn, updating its internal state
(which includes possibly modifying $T_k$) as it goes.
Suppose we are at the transition from tree $T_k$ to tree $T_{k+1}$,
which is done by first removing a set of edges $O$
and then adding another set of edges $I$.
$O$ defines a sub-forest $F_O$ of $T_k$,
and $I$ defines a sub-forest $F_I$ of $T_{k+1}$.
The key step in the algorithm is to determine whether the pair of paths
that terminate in a given node in two adjacent trees are mergeable,
\algorithmref{alg:mergeable}, which works as follows.
If a pair of paths is mergeable,
then the edges of the two paths must lie in $O$ and $I$, respectively.
Suppose an edge in $O$ has child $c$.
To see if $c$ is the base of a pair of mergeable paths,
the algorithm traverses up from $c$ in both $F_O$ and $F_I$;
terminating if a node in the other tree is found
(i.e., if the node traversed in $F_O$ is in $T_{k+1}$ or if the node traversed in $F_I$ is in $T_k$)
or if a pair of traversed nodes have the same time.
If these two traversals end in the same node $p$, the paths are mergeable.
Iterating over all edges in $O$ will thus find all mergeable pairs of paths.
There is often more than one pair of mergeable paths in a tree transition;
so, the algorithm merges pairs of mergeable paths,
starting with pairs that add the smallest number of new edges,
until no more are found.

\algorithmref{alg:extend} simplifies the algorithm in several ways --
for instance, the bookeeping required to keep track of $T_k$ and $T_{k+1}$ is omitted.
Furthermore, as described the algorithm does one left-to-right pass over the tree sequence;
in practice we do repeated passes in both directions until no changes can be made.
The main step that is omitted is a description of the \texttt{merge} operation,
which performs the actual extending of haplotypes.
This algorithm is essentially the same as \texttt{mergeable} in \algorithmref{alg:mergeable},
except with additional bookkeeping.
Roughly speaking, the algorithm
traverses up from the shared base node $c$,
doing the appropriate operations to insert the nodes along the path found in $T_k$
into the path in $T_{k+1}$.
To do this, some edges that end at $a_k$ will be extended to end at $a_{k+1}$;
some edges that begin at $a_k$ will be postponed to begin at $a_{k+1}$,
and some entirely new edges may be added, see \figref{fig:extending_diagram}.
Furthermore, the trees $T_k$ and $T_{k+1}$ (and corresponding forests $F_O$ and $F_I$)
need to be updated.

@@

\comment{Rename Dissimilarity and Similarity}

@@

%\comment{Delete Discrepancy Algorithm!!}
% 
% \begin{algorithm2e}[!ht]
% \SetStartEndCondition{ }{}{}%
% \SetKwProg{Fn}{def}{\string:}{}
% \SetKwFunction{Range}{range}%%
% \SetKw{KwTo}{in}
% \SetKwFor{For}{for}{\string:}{}%
% \SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
% \SetKwFor{While}{while}{:}{fintq}%
% \SetKw{Break}{break}
% \newcommand\forcond{$i$ \KwTo\Range{$n$}}
% \SetKwComment{Comment}{}
% \AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
% \DontPrintSemicolon\LinesNotNumbered
% 
% \caption{Discrepancy Function}\label{alg:disc}
% \input{$(T_1,T_2)$}
% \output{tree discrepancy, root-mean-square error}
% 
% \Fn{Shared Spans($T_1,T_2$)}{
% \For{pairs $(i,j)\in N_{T_1}\times N_{T_2}$}{
% 	$i = (t_i, L_i)$ and $j = (t_j, L_j)$ where $L_i, L_j\subset \left[0,L\right).\quad$ \Comment{Each node is a tuple (time, genome location).}
% 	\If{$(i,j)$ has equal sample sets, $s(i)=s(j)$}{
% 		Define $a_{ij}=|L_i\cap L_j|$.
% 	}
% 	\Else{Define $a_{ij}=0$.}
% Define matrix $A=\left[a_{ij}\right]\in M_{N_{T_1}\times N_{T_2}}(\R)$.\;
% }
% \KwRet{$A$}
% }
%\BlankLine
%\Fn{Discrepancy ($T_1,T_2$)}{
%\For{each node $i\in N_{T_1}$}{
%	Find nodes in $N_{T_2}$ which have the highest matching span.\;
%	$\operatorname{matches}_i = \argmax_{k\in N_{T_2}} a_{ik}.$\;
%	Let $j = \argmax_{k\in \operatorname{matches}_i}\frac{1}{1+|t_i-t_k|}$.\;
%	Define $\alpha_i = a_{ij}$.\;
%	\BlankLine
%}
%$$tree\ discrepancy = 1 - \frac{\sum_{i\in N_{T_1}} \alpha_i}{\operatorname{tr}\left(Shared Spans(T_1, T_1)\right)}$$\;
%\BlankLine\BlankLine
%$$rmse = \sqrt{\frac{\sum_{i\in N_{T_1}}|t_i - t_j|\alpha_i}{\operatorname{tr}\left(Shared Spans(T_1,T_1)\right)}}$$\;
%\BlankLine
%$$true\ proportion = \frac{\sum_{i\in N_{T_1}} \alpha_i}{\operatorname(tr)\left( Shared Spans(T_2, T_2)\right)}$$\;
%\BlankLine\BlankLine
%\KwRet{tree discrepancy, rmse, true proportion}
%}
%\end{algorithm2e}

@@

\begin{figure}[bht]
	\includegraphics[width=0.9\linewidth]{newplots_wo_ee/node_spans_wo_trapped_unary_regions_CDF.pdf}
	\caption{
        Node spans of simulated tree sequence (blue), simplified tree sequence (green), and extended tree sequence (red).
        \comment{TODO: nodes should be in the same order in all three lines,
        using the ordering from the true tree sequence.}
        \comment{TODO: add plot with percent wrongly added by node?}
    }
	\label{fig:node-spans}
\end{figure}

@@

\begin{figure}
	\includegraphics{dissimilarity.pdf}
\end{figure}

@@

\begin{tabular}[c|c|c|c|c|]
	Initial & Iteration 1 & Iteration 2 & Iteration 3 & Iteration 4 \\
	113463	&	76301	&	76059	&	76057	&	76057 \\
	113266 & 76084	&	75835	&	75833	&	75833 \\
	114489 & 76703 & 76436 & 76434 & 76434 \\
	114086	& 76550	& 76294	& 76291	& 76291
\end{tabular}

%\begin{algorithm2e}[!ht]
%	\SetStartEndCondition{ }{}{}%
%	\SetKwProg{Fn}{def}{\string:}{}
%	\SetKwFunction{Range}{range}%%
%	\SetKw{KwTo}{in}
%	\SetKwFor{For}{for}{\string:}{}%
%	\SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
%	\SetKwFor{While}{while}{:}{fintq}%
%	\SetKw{Break}{break}
%	\newcommand\forcond{$i$ \KwTo\Range{$n$}}
%	\AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
%	
%	\caption{Extends haplotypes across a tree sequence $TS$.
%    \comment{I think we can remove this now?}}
%	\label{alg:hap}
%	
%	\SetKwFunction{ex}{Extend Paths}
%	\SetKwFunction{fp}{Find Paths}
%	\DontPrintSemicolon
%	\LinesNotNumbered
%	%\SetAlgoNoEnd
%	\input{A tree sequence $\T$}
%	\output{New tree sequence,$\T'$}
%	\BlankLine
%	\Fn{\fp (tree sequence $\T$, forwards = \tn{True})}{
%	{
%		\eIf{forwards=\tn{True}}{$\T = \left(T_1,...,T_{|\T|}\right)$\BlankLine}
%		{\BlankLine
%			$\T = \left(T_{|\T|},..., T_1\right)$\;
%		}
%	}
%		\For{pairs $(T_k, T_{k+1})\in \T\times\T$}{
%		Compute subforests $F_O\subseteq T_k$ and $F_I\subseteq\in T_{k+1}$.\;
%		\For{edges $e$ in $F_O$}{
%			Let $c$ be child node.\;
%			$p_c^{F_O} = [c]$\;
%			$p_c^{F_I} = [c]$\;
%			$\bar{p}_c = []$ Path to be inserted\;
%			\While{$\exists f\in F_O$ such that $\tn{child}(f)=c$ and $\tn{deg}(\tn{parent}(f))=0$}{
%				$p = \tn{parent}(f)$\;
%				Append $p$ to $p_c^{F_O}$\;
%				$c = p$\;
%			}
%			\While{$\exists f\in F_I$ such that $\tn{child}(f)=c$ and $\tn{deg}(\tn{parent}(f))<3$}{
%				$p =\tn{parent}(f)$\;
%				Append $p$ to $p_c^{F_I}$\;
%				$c = p$\;
%			}
%			\If{($p_c^{F_O}\cap p_c^{F_I})\backslash\{c\}\neq\emptyset$}{
%				Merge paths up to intersection point\;
%				$\bar{p}_c = \hat{p}_c^{F_I} \cup \hat{p}^{F_I}_c$\;
%				and sort by node time.\;
%			}
%			Insert $\bar{p}_c$ into $T_{k+1}$\;		
%		}
%	}\;
%	Update $\T$.\;
%	\KwRet{$\T$}\;
%}
%\BlankLine						
%\Fn{\ex ($\T$)}{
%	$\T' \gets $ \fp($\T$, forwards =\tn{True})\;
%	$\T' \gets $ \fp($\T$, forwards = \tn{False})\;
%	\While{Number of edges in $\T$ $\neq$ Number of edges in $\T'$}{
%		$\T' \gets $ \fp($\T$, fowards=\tn{True})\;
%		$\T' \gets $ \fp($\T$, fowards=\tn{False})\;
%		
%	}
%	\KwRet{$\T'$}\;
%}
%\end{algorithm2e}
%
%% Can most likely delete the extend-edges algorithm %%
%\begin{algorithm2e}[!ht] 
%	
%\SetStartEndCondition{ }{}{}%
%\SetKwProg{Fn}{def}{\string:}{}
%\SetKwFunction{Range}{range}%%
%\SetKw{KwTo}{in}
%\SetKwFor{For}{for}{\string:}{}%
%\SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
%\SetKwFor{While}{while}{:}{fintq}%
%\SetKw{Break}{break}
%\newcommand\forcond{$i$ \KwTo\Range{$n$}}
%\AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
%
%\caption{Extends edges across a tree sequence $TS$. 
%\comment{Can be removed, probably -- unless this seems a lot simpler?}}
%\label{alg:edge}
%
%\SetKwFunction{ex}{Extend Edges}
%\SetKwFunction{fwd}{Extend}
%\DontPrintSemicolon
%\LinesNotNumbered
%%\SetAlgoNoEnd
%\input{A tree sequence $\T$}
%\output{New tree sequence,$\T'$}
%
%\BlankLine 
%
%\Fn{\fwd (tree sequence $\T$, forwards = \tn{True})}{
%{	
%\eIf{forwards=\tn{True}}{$\T = \left(T_1,...,T_{|\T|}\right)$\BlankLine}
%{\BlankLine
%	$\T = \left(T_{|\T|},..., T_1\right)$\;}
%}
%	\For{$T_k\in \T$}{
%	Find all edge pairs $(e_1,e_2)$ such that $e_1 = a\to b$ and $e_2 = b\to c$ for some nodes $a,b,c$.\;
%	\For{edges $f$ in $T_{k+1}$}{
%		\If{$f = a \to c$ and $b$ is not in $T_{k+1}$}{
%		Remove $f$ from $T_{k+1}$.\;
%		Add $e_1$ and $e_2$ to $T_{k+1}$.\;
%		Update $\T$.\;
%	}
%	}
%}\;						
%\KwRet{$\T$}\;
%}
%\BlankLine						
%\Fn{\ex ($\T$, iter=100)}{
%	\For{$*$ \KwTo \Range(iter)}{
%		$\T' \gets \T$.\fwd(fowards=\tn{True})\;
%		$\T' \gets \T$.\fwd(fowards=\tn{False})\;
%		\If{Number of edges in $\T$ = Number of edges in $\T'$}{
%			\Break\BlankLine}
%		}
%}
%\KwRet{$\T'$}\;
%\end{algorithm2e}
